

# Reproducible Research: Peer Assessment 1

---

title: "Reproducible Research: Peer Assessment 1"
output:
        html document:
                keep_md: TRUE

---

Here is the code for reading in and processing the data

```{r Processing Data Code}
library(dplyr)
library(tidyverse)
library(reshape2)
library(lubridate)
library(ggplot2)

activdata <-read.csv("activity.csv")

head(activdata)

activdata <- tbl_df(activdata)
activdata
summary(activdata$steps)

activdata$dayname <- wday(activdata$date, label = TRUE)

activdata$day <- day(activdata$date)

activdata$month <- month(activdata$date)
```

Now plotting the total number of steps per day and calculating the mean and median number of total number of steps per day

```{r Mean total number of steps taken per day}

stepsday <- tapply(activdata$steps, activdata$date, sum, na.rm = TRUE)
qplot( x = stepsday, geom = "histogram") + geom_histogram(color = "white") + theme_bw() +
        labs(x = "Steps taken per day", y = "Frequency")

meansteps <- mean(activdata$steps, na.rm = T)
mediansteps <- median(activdata$steps, na.rm = T)
print("Mean total number of steps per day:")
meansteps
print("Median total number of steps per day:")
mediansteps

```

Writing code to plot average number of steps for all intervals,averaged across all days and
printing the interval with maximum average number of steps

```{r Relationship between intervals and number of steps taken}

intervalav <- tapply(activdata$steps, activdata$interval, mean, na.rm = TRUE)
intervalav <- cbind(rownames(intervalav), intervalav)
intervalav <- tbl_df(intervalav)
colnames(intervalav) <- c("interval", "average")
intervalav <- tbl_df(intervalav)
intervalav$interval <- as.numeric(intervalav$interval)
intervalav$average <- as.numeric(intervalav$average)
qplot(data = intervalav, x = interval, y = average, geom = c("point","line")) +
        theme_bw() + labs(title = "Time series plot of average steps per interval of time", x = "Intervals",
                          y = "Average number of steps") 


maxinterval <- filter(intervalav, average == max(average))
print("Interval with maximum average number of steps is: ")
maxinterval$interval
print("The maximun average number of steps are: ")
maxinterval$average

```

Writing code for reporting NA's, filling in the NA values and drawing means and medians again

```{r Imputing Missing Values}

summary(activdata)

print("Looking at the summary we conclude that the data has 2304 rows with NA values")

# now replacing the NA values 


intervalav <- tapply(activdata$steps, activdata$interval, mean, na.rm = TRUE)

for(i in 1:length(activdata$steps)){
        
        if (is.na(activdata$steps[i])){
                
                activdata$steps[i] <- intervalav[as.character(activdata$interval[i])]
        }
        
}

# the activdata dataset has been modified to fill in the na values with the means

print("The dataset has been modified with values put in place of NA")

stepsday <- tapply(activdata$steps, activdata$date, sum, na.rm = TRUE)
qplot( x = stepsday, geom = "histogram") + geom_histogram(color = "white") + theme_bw() +
        labs(x = "Steps taken per day", y = "Frequency")

meansteps <- mean(activdata$steps, na.rm = T)
mediansteps <- median(activdata$steps, na.rm = T)
print("Mean total number of steps per day:")
meansteps
print("Median total number of steps per day:")
mediansteps

```

Finally making the plots for weekends and weekdays 

```{r Weekend and Weekday plots with updated data}

activdata$daytype <- factor(activdata$dayname %in% c("Sat", "Sun"), labels = c("weekday","weekend"))
weekdata <- tapply(activdata$steps, list(activdata$interval, activdata$daytype), mean, na.rm = TRUE)
weekdata <- cbind(rownames(weekdata), weekdata[,1], weekdata[,2])
weekdata1 <- tbl_df(weekdata)
weekdata1$interval <- as.numeric(rownames(weekdata))
weekdata <- melt(weekdata1, id = "interval")
g <- ggplot(data = weekdata, aes(x = interval, y = value)) + facet_grid(variable~.) + theme_bw()
g + geom_point() + geom_line() + labs(x = "Intervals", y = "Average number of steps", title = "Average number of steps on weekends and weekdays")


```
